---
title: "Introduction to bvartools"
author: "Franz X. Mohr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to bvartools}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The package `bvartools` implements some common functions used for Bayesian inference for mulitvariate time series models. It should give researchers maximum freedom in setting up an MCMC algorithm in R and keep calculation time limited at the same time. This is achieved by implementing posterior simulation functions in C++. Its main features are

-   Posterior simulation functions, which are written in C++ for faster computation
-   The `bvar` and `bvec` function collects the output of a Gibbs sampler in standardised objects, which can be used for further analyses
-   Further functions such as `predict`, `irf`, `fevd` for forecasting, impulse response analysis and forecast error variance decomposition, respectively.

This vignette provides the code to set up and estimate a basic Bayesian VAR (BVAR) model with the `bvartools` package. For this illustration the dataset E1 from L端tkepohl (2007) is used. It contains data on West German fixed investment, disposable income and consumption expenditures in billions of DM from 1960Q1 to 1982Q4.

```{r data, fig.align='center', fig.height=5, fig.width=4.5}
library(bvartools)

data("e1")
e1 <- diff(log(e1))

plot(e1) # Plot the series
```

The `gen_var` function produces the inputs `y` and `x` for the BVAR estimator, where `y` is a matrix of dependent variables and `x` is the matrix of regressors for the model

$$y_t = A x_t + u_t,$$
with $u_t \sim N(0, \Sigma)$.

```{r}
data <- gen_var(e1, p = 2, deterministic = "const")

y <- data$Y[, 1:73]
x <- data$Z[, 1:73]
```

As in L端tkepohl (2007) only the first 73 observations are used.

## Estimation

### Frequentist estimator

We calculate frequentist VAR estimates using the standard formula $y x' (x x')^{-1}$ to obtain a benchmark for the Bayesian estimator. The parameters are obtained by OLS:

```{r}
A_freq <- tcrossprod(y, x) %*% solve(tcrossprod(x)) # Calculate estimates
round(A_freq, 3) # Round estimates and print
```

And $\Sigma$ is calculated by

```{r}
u_freq <- y - A_freq %*% x
u_sigma_freq <- tcrossprod(u_freq) / (ncol(y) - nrow(x))
round(u_sigma_freq * 10^4, 2)
```

These are the same values as in L端tkepohl (2007).

### Bayesian estimator

The following code is a Gibbs sampler for a simple VAR model with non-informative priors.

```{r flat prior}
# Reset random number generator for reproducibility
set.seed(1234567)

iter <- 30000 # Number of iterations of the Gibbs sampler
burnin <- 10000 # Number of burn-in draws
store <- iter - burnin

t <- ncol(y) # Number of observations
k <- nrow(y) # Number of endogenous variables
m <- k * nrow(x) # Number of estimated coefficients

# Set (uninformative) priors
a_mu_prior <- matrix(0, m) # Vector of prior parameter means
a_v_i_prior <- diag(0, m) # Inverse of the prior covariance matrix

u_sigma_df_prior <- 0 # Prior degrees of freedom
u_sigma_scale_prior <- diag(0, k) # Prior covariance matrix
u_sigma_df_post <- t + u_sigma_df_prior # Posterior degrees of freedom

# Initial values
u_sigma_i <- diag(.00001, k)
u_sigma <- solve(u_sigma_i)

# Data containers for posterior draws
draws_a <- matrix(NA, m, store)
draws_sigma <- matrix(NA, k^2, store)

# Start Gibbs sampler
for (draw in 1:iter) {
  # Draw conditional mean parameters
  a <- post_normal(y, x, u_sigma_i, a_mu_prior, a_v_i_prior)
  
  # Draw variance-covariance matrix
  u <- y - matrix(a, k) %*% x # Obtain residuals
  u_sigma_scale_post <- solve(u_sigma_scale_prior + tcrossprod(u))
  u_sigma_i <- matrix(rWishart(1, u_sigma_df_post, u_sigma_scale_post)[,, 1], k)
  u_sigma <- solve(u_sigma_i) # Invert Sigma_i to obtain Sigma
  
  # Store draws
  if (draw > burnin) {
    draws_a[, draw - burnin] <- a
    draws_sigma[, draw - burnin] <- u_sigma
  }
}
```

After the Gibbs sampler has finished, point estimates can be obtained as the mean of the posterior draws:

```{r}
A <- rowMeans(draws_a) # Obtain means for every row
A <- matrix(A, k) # Transform mean vector into a matrix
A <- round(A, 3) # Round values
dimnames(A) <- list(dimnames(y)[[1]], dimnames(x)[[1]]) # Rename matrix dimensions

A # Print
```

```{r}
Sigma <- rowMeans(draws_sigma) # Obtain means for every row
Sigma <- matrix(Sigma, k) # Transform mean vector into a matrix
Sigma <- round(Sigma * 10^4, 2) # Round values
dimnames(Sigma) <- list(dimnames(y)[[1]], dimnames(y)[[1]]) # Rename matrix dimensions

Sigma # Print
```

The means of the coefficient draws are very close to the results of the frequentist estimatior, which would be expected with non-informative priors.

## `bvar` objects

The `bvar` function can be used to collect relevant output of the Gibbs sampler into a standardised object, which can be used by further functions such as `predict` to obtain forecasts or `irf` for impulse respons analysis.

```{r bvar-object}
bvar_est <- bvar(y = y, x = x, A = draws_a[1:18,],
                 C = draws_a[19:21, ], Sigma = draws_sigma)
```

## Forecasts

Forecasts with credible bands can be obtained with the function `predict`. If the model contains deterministic terms, new values can be provided in the argument `new_D`. If no values are provided, the function sets them to zero. The number of rows of `new_D` must be the same as the argument `n.ahead`.

```{r forecasts, fig.width=5.5, fig.height=5.5}
bvar_pred <- predict(bvar_est, n.ahead = 10, new_D = rep(1, 10))

plot(bvar_pred)
```

## Impulse response analysis

Currently, `bvartools` supports forecast error, orthogonalised, and generalised impulse response functions.

### Forecast error impulse response

```{r feir, fig.width=5.5, fig.height=4.5}
IR <- irf(bvar_est, impulse = "income", response = "cons", n.ahead = 8)

plot(IR, main = "Forecast Error Impulse Response", xlab = "Period", ylab = "Response")
```

### Orthogonalised impulse response

```{r oir, fig.width=5.5, fig.height=4.5}
OIR <- irf(bvar_est, impulse = "income", response = "cons", n.ahead = 8, type = "oir")

plot(OIR, main = "Orthogonalised Impulse Response", xlab = "Period", ylab = "Response")
```

### Generalised impulse response

```{r gir, fig.width=5.5, fig.height=4.5}
GIR <- irf(bvar_est, impulse = "income", response = "cons", n.ahead = 8, type = "gir")

plot(GIR, main = "Generalised Impulse Response", xlab = "Period", ylab = "Response")
```

## Forecast error variance decomposition

Default forecast error variance decomposition (FEVD) is based on orthogonalised impulse responses (OIR).

```{r fevd-oir, fig.width=5.5, fig.height=4.5}
bvar_fevd_oir <- fevd(bvar_est, response = "cons")

plot(bvar_fevd_oir, main = "OIR-based FEVD of consumption")
```

It is also possible to calculate FEVDs, which are based on generalised impulse responses (GIR). Note that these do not automatically add up to unity.

```{r fevd-gir, fig.width=5.5, fig.height=4.5}
bvar_fevd_gir <- fevd(bvar_est, response = "cons", type = "gir")

plot(bvar_fevd_gir, main = "GIR-based FEVD of consumption")
```


## References

Koop, G., Pesaran, M. H., & Potter, S.M. (1996). Impulse response analysis in nonlinear multivariate models. *Journal of Econometrics 74*(1), 119-147.

L端tkepohl, H. (2007). *New introduction to multiple time series analyis*. Berlin: Springer.

Pesaran, H. H., & Shin, Y. (1998). Generalized impulse response analysis in linear multivariate models. *Economics Letters, 58*(1), 17-29.