---
title: "Horse Races"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Horse Races}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(bvartools)
```

## Introduction


## Workflow

```{r}
data("us_macrodata")

#us_macrodata <- diff(us_macrodata)

expanding_window_start <- 2006
```


### Producing candidate models

```{r}
# Shared specifications
iterations <- 2000
burnin <- 1000
```


First, create a series of traditional VAR models:

```{r}
var_default <- create_bvarmodel(us_macrodata,
                                p = 1:4,
                                deterministic = "const",
                                iterations = iterations,
                                burnin = burnin)

var_default <- use_expanding_window(var_default, start = expanding_window_start)

var_default <- add_priors(var_default,
                          coef = list(v_i = 1 / 10, v_i_det = 1 / 100),
                          sigma = list(df = 3, scale = 1))

var_default <- add_initial_values(var_default)
```

Second, create a series of VAR models with stochastic search variable selection (SSVS):

```{r}
var_ssvs <- create_bvarmodel(us_macrodata,
                            p = 4,
                            varsel = "ssvs",
                            deterministic = "const",
                            iterations = iterations,
                            burnin = burnin)

var_ssvs <- use_expanding_window(var_ssvs, start = expanding_window_start)

var_ssvs <- add_priors(var_ssvs,
                          coef = list(v_i = 1 / 10, v_i_det = 1 / 100),
                          sigma = list(df = 3, scale = 1),
                          varsel = list(inprior = 0.5, tau = c(0.05, 10), exclude_deterministic = TRUE))

var_ssvs <- add_initial_values(var_ssvs)
```

Thrird, create a series of VAR models with Bayesian variable selection à la Korobilis (2013):

```{r}
var_bvs <- create_bvarmodel(us_macrodata,
                            p = 4,
                            varsel = "bvs",
                            deterministic = "const",
                            iterations = iterations,
                            burnin = burnin)

var_bvs <- use_expanding_window(var_bvs, start = expanding_window_start)

var_bvs <- add_priors(var_bvs,
                          coef = list(v_i = 1 / 10, v_i_det = 1 / 100),
                          sigma = list(df = 3, scale = 1),
                          varsel = list(inprior = 0.5, exclude_deterministic = TRUE))

var_bvs <- add_initial_values(var_bvs)
```

Use `combine_models` to make one object:

```{r}
models <- combine_models(var_default,
                         var_ssvs,
                         var_bvs)
```

<!-- Ensure that all models look at the same time series: -->

<!-- ```{r} -->
<!-- models <- align_model_obs(models) -->
<!-- ``` -->


### Draw posteriors

```{r}
models <- add_posterior_coefficients(models)
```


### Forecasting

Forecasts are obtained in two steps. First, function `add_forecast_input` generates the data of the forecast periods, where deterministic terms can be provided in the argument `deterministic` and unmodelled variables in the argument `exogen`. If they are not provided, the function tries to obtain them from the model and gives a message, if it does not succeed. Second, function `add_posterior_forecasts` then simulates the forecasts.

```{r}
models <- add_forecast_input(models, n_ahead = 4)
models <- add_posterior_forecasts(models)
```

Calculate forecast errors on which out-of-sample selection criteria will be based:

```{r}
models <- add_forecast_errors(models, test_sample = us_macrodata)
```


### Evaluation

Use function `add_posterior_loglik` to add log-likelihoods, on which in-sample selection criteria will be based:

```{r}
models <- add_posterior_loglik(models)
```

Calculate in-sample and out-of-sample selection criteria:

```{r}
sc <- selection_criteria(models)
```

Compare the selection criteria. For out-of-sample criteria, the first model is used as a reference for comparison:

```{r}
print(sc, relative = 1)
```


#### In-sample plots

```{r, fig.width=5.5, fig.height=4.5}
plot(sc, criterion = "BIC")
```

```{r, fig.width=5.5, fig.height=4.5}
plot(sc, criterion = "AIC")
```

#### Out-of-sample plots

```{r, fig.width=5.5, fig.height=4.5}
plot(sc, criterion = "FE")
```
```{r, fig.width=5.5, fig.height=4.5}
plot(sc, criterion = "AFE")
```

```{r, fig.width=5.5, fig.height=4.5}
plot_forecast_errors_by_period(models, criterion = "FE")
```


## References

Korobilis, D. (2013). VAR forecasting using Bayesian variable selection. *Journal of Applied Econometrics, 28*(2), 204-230. <https://doi.org/10.1002/jae.1271>