---
title: "Moving from bvartools 0.3.0 to 1.0.0"
author: "Franz X. Mohr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Moving from bvartools 0.3.0 to 1.0.0}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## What this release is for

Version 0.3.0 is a transition release. It is the functionality of bvartools as
it has been on CRAN, with a small number of correctness fixes, and it is the
last version before 1.0.0 reorganises the package around a different set of
functions.

Nothing in this release stops working. What it adds is a message: the first
time in a session that you use a function which 1.0.0 does not have any more,
bvartools says so and names what takes its place. Upgrading to this release is
therefore a way to find out what a later upgrade to 1.0.0 will cost you, while
everything still runs.

The messages are shown once per function per session. To switch them off:

```{r}
options(bvartools.transition.messages = FALSE)
```

## Renamed functions

These do the same thing under a new name. The arguments are unchanged.

| 0.3.0 | 1.0.0 |
|---|---|
| `gen_var()` | `create_bvarmodel()` |
| `gen_vec()` | `create_bvecmodel()` |
| `bvec_to_bvar()` | `vec_to_var()` |
| `kalman_dk()` | `kalman_durbin_koopman_2002()` |
| `stochvol_ksc1998()` | `stochvol_ksc_1998()` |
| `stochvol_ocsn2007()` | `stochvol_ocsn_2007()` |
| `stoch_vol()` | `stochvol_ksc_1998()` |
| `bvs()` | `post_bvs()` |

`stoch_vol()` was a wrapper for the algorithm of Kim, Shephard and Chib (1998),
which `stochvol_ksc_1998()` implements directly, so the wrapper has no separate
successor.

## A changed workflow for posterior simulation

`draw_posterior()`, `bvarpost()` and `bvecpost()` are replaced by a sequence of
functions, each of which adds one thing to the model object. Where 0.3.0 has

```{r}
object <- gen_var(data, p = 2, deterministic = "const")
object <- add_priors(object)
object <- draw_posterior(object)
```

1.0.0 has

```{r}
object <- create_bvarmodel(data, p = 2, deterministic = "const")
object <- add_priors(object)
object <- add_initial_values(object)
object <- add_posterior_coefficients(object)
```

with `add_posterior_forecasts()` and `add_posterior_loglik()` producing the
forecasts and the log likelihood that `draw_posterior()` used to produce in the
same call. Splitting them apart is what lets a model be estimated once and then
have forecasts added, or the log likelihood recomputed, without repeating the
simulation.

## Removed without a successor

Dynamic factor models are removed from bvartools in 1.0.0 entirely. That covers
`dfm()`, `dfmpost()` and `gen_dfm()`, the `add_priors()`, `plot()`, `summary()`
and `thin()` methods for objects of class `dfm`, and the example data set
`bem_dfmdata`. A data set cannot announce itself, so this is the only notice
`bem_dfmdata` gets.

`post_normal_covar_const()` and `post_normal_covar_tvp()` are also removed with
nothing taking their place.

## Renamed classes, and the methods that follow them

The model classes are renamed in 1.0.0. The generics are the same, so code that
calls `plot()`, `summary()`, `predict()` or `thin()` on a model object keeps
working; what changes is the name of the class those methods are written for,
which matters if you dispatch on it yourself or test for it with `inherits()`.

| 0.3.0 | 1.0.0 |
|---|---|
| `bvar` | `bvarmodel` |
| `bvec` | `bvecmodel` |
| `bvarlist` | `modellist` |

These do not produce a message, because the function you call is unchanged.

## Same name, different input

These functions exist in 1.0.0 under the same name but do not take the same
input, because they take the reorganised model object. They produce no message
either, for the same reason: the name is still there. Read their documentation
before assuming a call carries over.

`add_priors()`, `bvar()`, `bvec()`, `irf()`, `fevd()`, `inclusion_prior()`,
`minnesota_prior()`, `ssvs_prior()`.

## Fixes in this release

Four defects in the CRAN sources are fixed here. Two of them change results.

* **Stochastic volatility with an observation far out in the tails.**
  `stochvol_ksc1998()` and `stochvol_ocsn2007()` sampled the mixture indicator
  from weights formed as densities and normalised by their sum. Where an
  observation lies far enough out in the tails of every component, each density
  underflows to zero, the row sums to zero, the weights become `NaN`, and the
  indicator runs one past the last component -- which ends the call with
  `Mat::elem(): index out of bounds`. The weights are now formed in logs and
  shifted by their row maximum before they are exponentiated. This is
  algebraically the same calculation, and draws from a given seed are
  unchanged: verified bit for bit against the previous implementation. Only the
  case that used to fail behaves differently, and it now returns a draw.

* **Argument checking in the same two functions.** `sigma`, `h_init` and
  `constant` were indexed on trust, so a vector of the wrong length surfaced as
  `Mat::elem(): index out of bounds` rather than as a statement about the
  argument. Each is now checked against the number of columns of `y` and named
  in the error.

* **Impulse responses and variance decompositions of structural models.** A
  structural model keeps its contemporaneous block separately, so its
  coefficient draws are the structural `A_i` and its covariance draws the
  covariance of the structural errors. The forecast error, orthogonalised and
  generalised recursions want the reduced form. Given the structural quantities
  they returned numbers that belong to no model at all -- and the same numbers
  for all three types, since the structural error covariance makes the
  orthogonalisation degenerate. `irf()` with `type` of `"feir"`, `"oir"` or
  `"gir"`, and `fevd()` with `"oir"` or `"gir"`, now refuse a structural model
  and say why. Use `"sir"` or `"sgir"`, which are what a structural model is
  for and are unchanged, or estimate the model with `structural = FALSE`.

* **Seasonal terms for data of frequency one.** `gen_vec()` warned that no
  seasonal dummies are generated for a series of frequency one and then added
  them anyway, stopping with `object 'seas' not found`. It now does what the
  warning says. `gen_var()` was never affected.

## Getting the messages out of a script

Beyond the option above, the messages are ordinary conditions, so
`suppressMessages()` works on any single call:

```{r}
object <- suppressMessages(gen_var(data, p = 2, deterministic = "const"))
```

They are messages rather than warnings deliberately: they will not become
errors under `options(warn = 2)`, and they will not fail a check that treats
warnings as failures.
